---
title: "gtsummary in quarto"
format: 
  html:
    theme: quartz
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(gtsummary)
library(dplyr)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd",
               "id", "nsibs", "samp", "race_eth", "sex", "region",
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"),
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |>
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes")))
```

# Customization of `tbl_summary()`

*This table is a summary of the demographic characteristics of male and female participants in the United States.*

We’re using some data from the [**National Longitudinal Survey of Youth 1979**](https://www.nlsinfo.org/content/cohorts/nlsy79)**,** a cohort of American young adults aged 14-22 at enrollment in 1979. They continue to be followed to this day, and there is a wealth of publicly available data [**online**](https://www.nlsinfo.org/investigator/pages/login.jsp). I’ve downloaded the answers to a survey question about whether respondents wear glasses, a scale about their eyesight with glasses, their (NLSY-assigned 😒) race/ethnicity, their sex (presumably at birth, but unclear), their family’s income in 1979, and their age at the birth of their first child.

The median age at birth for males was 25 and 22 for females. (@tbl-agebir)

```{r}
#| label: tbl-agebir
#| tbl-cap: "Descriptive statistics characterizing age at birth"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

Majority of males had either excellent eyesight or wear glasses, wheras majority of females wear glasses. (@tbl-eyesight)

```{r}
#| label: tbl-eyesight
#| tbl-cap: "Descriptive statistics characterizing eyesight"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

There were a lot of non-black,non-Hispanic males and females who slept (@tbl-sex)

```{r}
#| label: tbl-sex
#| tbl-cap: "Descriptive statistics stratified by sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

There were a lot of people who slept (@tbl-sleep)

```{r}
#| label: tbl-sleep
#| tbl-cap: "Descriptive statistics including sleep"

tbl_sleep<- tbl_summary(
  nlsy,
  by = region_cat,
  include = c(sleep_wknd,eyesight_cat,
              race_eth_cat,income,nsibs), label = list(
    sleep_wknd ~ "Sleep on Weekends",
    eyesight_cat ~ "Eyesight",
    race_eth_cat ~ "Race/ethnicity",
    income ~ "Income",
    nsibs ~ "Number of siblings"
    )) |>
  add_overall()

tbl_sleep
```

```{r}
stats <- list(n = nrow(data),
              median_sleep_wknd = median(nlsy$sleep_wknd, na.rm = TRUE))
stats
inline_text(tbl_sleep,variable="sleep_wknd",column= "Northeast", pattern = "{median}")
```

```{r}
median_sleep <- median(nlsy$sleep_wknd, na.rm = TRUE)
```

The median number of hours of sleep was `r median_sleep`

```{r}
# tbl_sleep$table_body
inline_text(tbl_sleep, variable ="sleep_wknd", 
            column = "stat_0")
inline_text(tbl_sleep,variable = "eyesight_cat",
            level= "Excellent",
            column = "stat_0")
```

In the sample the average sleep on weekends was `r inline_text(tbl_sleep, variable = "sleep_wknd", column = "stat_0")`.

```{r}

mean_sleep_wknd <- mean(nlsy$sleep_wknd, na.rm = TRUE)

inline_text(tbl_sleep,variable="sleep_wknd",column= "Northeast", pattern = "{stats$mean_sleep_wknd}")
```

In the sample the mean sleep on weeknds was `r mean(nlsy$sleep_wknd, na.rm = TRUE)` .

```{r}
#| label: tbl-sleep-means
#| tbl-cap: "Descriptive statistics including sleep"

tbl_sleep2<- tbl_summary(
  nlsy,
  by = region_cat,
  include = c(sleep_wknd,eyesight_cat,
              race_eth_cat,income,nsibs), label = list(
    sleep_wknd ~ "Sleep on Weekends",
    eyesight_cat ~ "Eyesight",
    race_eth_cat ~ "Race/ethnicity",
    income ~ "Income",
    nsibs ~ "Number of siblings"
    ),
  statistic = list(all_continuous()~"{mean},sd ={sd}")) |>
  add_overall()
tbl_sleep2
```

```{r}
inline_text(tbl_sleep2,variable ="sleep_wknd",
            column="Northeast")
inline_text(tbl_sleep2, variable = "sleep_wknd", 
            column = "Northeast", pattern = "{mean}")
```

There are multiple ways we could report the median, `r median_sleep`, or the mean `r mean_sleep_wknd`.
